A follow-up on intermediate-mass black hole candidates in the second LIGO--Virgo observing run with the Bayes Coherence Ratio

### 2022-08-03 main.tex
Final version

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6\ M_\odot$) will provide clues to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. 
Ground-based gravitational-wave detectors are sensitive to mergers that can form intermediate-mass black holes weighing up to $\sim 450\ M_\odot$.
However, ground-based detector data contain numerous incoherent short duration noise transients that can mimic the gravitational-wave signals from merging intermediate-mass black holes, limiting the sensitivity of searches. 
Here we \green{follow-up on} binary black hole merger \green{candidates} using a ranking statistic that measures the coherence or incoherence of triggers in multiple-detector data. 
We use this statistic to \green{rank} candidate events,  \green{initially identified by all-sky search pipelines,} with lab-frame total masses $\gtrsim55\ M_\odot$ using data from LIGO's second observing run. 
Our analysis does not yield evidence for new intermediate-mass black holes. 
However, we find support for eight stellar-mass binary black holes not reported in the first LIGO-Virgo gravitational wave transient catalog GWTC-1, seven of which have been previously reported by other catalogs.
\end{abstract}

@@

Since it is impossible to shield ground-based gravitational-wave detectors from gravitational-wave signals, the LVK empirically estimates the background by repeatedly time-shifting strain data by amounts larger than the light-travel time between the two LIGO detectors~\citep{GWTC1}. We use time-shifted data to generate $\bgrdbcr$, \green{the background ranking statistic}. Following this, we calculate the \green{fraction of $\bgrdbcr$ greater than or equal to a $\candbcr$, the candidate ranking statistic}:
\begin{equation}
    p_1^b = \frac{\text{Count of } \bgrdbcr \leq \candbcr}{\text{Count of } \bgrdbcr} \ .
\end{equation}
Given a set of simulated signals \green{and their ranking statistic $\injbcr$, one may calculate the fraction of $\injbcr$ greater than or equal to a $\candbcr$}: 
\begin{equation}
    p_1^s = \frac{\text{Count of } \injbcr \leq \candbcr}{\text{Count of } \injbcr} \ .
\end{equation}
\green{With $p_1^b$ and $p_1^s$ it is possible to compute a candidate’s 
\pastro, the probability that a candidate is of astrophysical origin:}
\begin{equation}
    \pastro = \frac{p_1^s}{p_1^s + p_1^b}\ .
\end{equation}
However, for this study we do not have an astrophysical distribution of simulated signals and so we cannot compute $p_1^s$ \green{or consequently $\pastro$}. 
Instead we opt for a frequentist $p$-value probability that a candidate is inconsistent with the background. 
As we have $k$ candidates, each with a \candbcr, we calculate a false-alarm probability \fap that accounts for trial factors given by
\begin{equation}
    \fap = 1 - (1-p_1^b)^k \ .
\end{equation}
Finally, we compute the probability that a candidate is inconsistent with the background:
\begin{equation}
\pastrobcr = 1 - \fap \ . 
\end{equation}
When $\pastrobcr \ll 1$, the event is consistent with the background distribution. 
Conversely, when $\pastrobcr \approx 1$ the event is inconsistent with the background distribution, and is therefore a promising gravitational-wave candidate.

\green{It is important to note that 
$\pastrobcr$ (the probability that an event is not part of the background distribution) is not the same as $\pastro$, which requires an astrophysical set of simulated signals.}

@@

To generate the PSD, we take 31 neighboring off-source non-overlapping  4-second  segments of time-series data before the analysis data segment $d_i$. A Tukey window with a 0.2-second roll-off is applied to each data segment to suppress spectral leakage. After this, we fast-Fourier transform and median-average the segments to create a PSD~\citep{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\citep{psd_student_t, chatziioannou2019noise, Biscoveanu:2020:PhRvD}. To marginalize over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step. \green{This post-processing step reduces the percentage of background $\bgrdbcr>0$ by $\sim49\%$.} The details of this calculation are in the Appendix~\ref{sec:psd-marginalization}.

@@

All analyses (inclusive of test and failed analyses) performed for this study used ${1.3\mathrm{M}}$ core-hours, amounting to a carbon footprint of ${\sim1.67\ \mathrm{t}}$ of ${\text{CO}_2}$ (using the U.S. average electricity source emissions of ${0.429\ \text{kg/kWh}}$~\citep{greenhouse} and ${0.3\ \text{kWh}}$ for each CPU).

@@

\section{Marginalizing over PSD statistical uncertainties}\label{sec:psd-marginalization}
To generate the results presented in Table~\ref{tab:results}, we applied a post-processing step to marginalize the uncertainty in the PSD. In Fig.~\ref{fig:bcrCdf}, we demonstrate the impact of the post-processing step. Marginalizing over uncertainty in the PSD yields an improvement in the separation of the noise and signal distributions (left plot). Quantitatively, at a threshold $\bcr^T=0$ the post-processing step \green{reduces the percentage of} background $\bcr > \bcr^T$ from $60\%$ to $25\%$ \green{(a $58\%$ improvement)} in the August 13 - 21, 2017 time-frame of data. \green{For the entirety of O2, PSD marginalization reduces the percentage of $\bcr > \bcr^T$ from $64\%$ to $33\%$ (a $\sim49\%$ improvement). }

>>>>

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6\ M_\odot$) will provide clues to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. 
Ground-based gravitational-wave detectors are sensitive to mergers that can form intermediate-mass black holes weighing up to $\sim 450\ M_\odot$.
However, ground-based detector data contain numerous incoherent short duration noise transients that can mimic the gravitational-wave signals from merging intermediate-mass black holes, limiting the sensitivity of searches. 
Here we follow-up on binary black hole merger candidates using a ranking statistic that measures the coherence or incoherence of triggers in multiple-detector data. 
We use this statistic to rank candidate events,  initially identified by all-sky search pipelines, with lab-frame total masses $\gtrsim55\ M_\odot$ using data from LIGO's second observing run. 
Our analysis does not yield evidence for new intermediate-mass black holes. 
However, we find support for eight stellar-mass binary black holes not reported in the first LIGO-Virgo gravitational wave transient catalog GWTC-1, seven of which have been previously reported by other catalogs.
\end{abstract}

@@

Since it is impossible to shield ground-based gravitational-wave detectors from gravitational-wave signals, the LVK empirically estimates the background by repeatedly time-shifting strain data by amounts larger than the light-travel time between the two LIGO detectors~\citep{GWTC1}. We use time-shifted data to generate $\bgrdbcr$, the background ranking statistic. Following this, we calculate the fraction of $\bgrdbcr$ greater than or equal to a $\candbcr$, the candidate ranking statistic:
\begin{equation}
    p_1^b = \frac{\text{Count of } \bgrdbcr \leq \candbcr}{\text{Count of } \bgrdbcr} \ .
\end{equation}
Given a set of simulated signals and their ranking statistic $\injbcr$, one may calculate the fraction of $\injbcr$ greater than or equal to a $\candbcr$: 
\begin{equation}
    p_1^s = \frac{\text{Count of } \injbcr \leq \candbcr}{\text{Count of } \injbcr} \ .
\end{equation}
With $p_1^b$ and $p_1^s$ it is possible to compute a candidate’s 
\pastro, the probability that a candidate is of astrophysical origin:
\begin{equation}
    \pastro = \frac{p_1^s}{p_1^s + p_1^b}\ .
\end{equation}
However, for this study we do not have an astrophysical distribution of simulated signals and so we cannot compute $p_1^s$ or consequently $\pastro$. 
Instead we opt for a frequentist $p$-value probability that a candidate is inconsistent with the background. 
As we have $k$ candidates, each with a \candbcr, we calculate a false-alarm probability \fap that accounts for trial factors given by
\begin{equation}
    \fap = 1 - (1-p_1^b)^k \ .
\end{equation}
Finally, we compute the probability that a candidate is inconsistent with the background:
\begin{equation}
\pastrobcr = 1 - \fap \ . 
\end{equation}
When $\pastrobcr \ll 1$, the event is consistent with the background distribution. 
Conversely, when $\pastrobcr \approx 1$ the event is inconsistent with the background distribution, and is therefore a promising gravitational-wave candidate.

It is important to note that 
$\pastrobcr$ (the probability that an event is not part of the background distribution) is not the same as $\pastro$, which requires an astrophysical set of simulated signals.

@@

To generate the PSD, we take 31 neighboring off-source non-overlapping  4-second  segments of time-series data before the analysis data segment $d_i$. A Tukey window with a 0.2-second roll-off is applied to each data segment to suppress spectral leakage. After this, we fast-Fourier transform and median-average the segments to create a PSD~\citep{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\citep{psd_student_t, chatziioannou2019noise, Biscoveanu:2020:PhRvD}. To marginalize over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step. 
This post-processing step reduces the percentage of background $\bgrdbcr>0$ by $\sim49\%$.
The details of this calculation are in the Appendix~\ref{sec:psd-marginalization}.

@@

All analyses (inclusive of test and failed analyses) performed for this study used ${0.6\mathrm{M}}$ core-hours, amounting to a carbon footprint of ${\sim77\ \mathrm{t}}$ of ${\text{CO}_2}$ (using the U.S. average electricity source emissions of ${0.371\ \text{kg/kWh}}$~\citep{greenhouse} and ${0.3\ \text{kWh}}$ for each CPU).

@@

\section{Marginalizing over PSD statistical uncertainties}\label{sec:psd-marginalization}
To generate the results presented in Table~\ref{tab:results}, we applied a post-processing step to marginalize the uncertainty in the PSD. In Fig.~\ref{fig:bcrCdf}, we demonstrate the impact of the post-processing step. Marginalizing over uncertainty in the PSD yields an improvement in the separation of the noise and signal distributions (left plot). Quantitatively, at a threshold $\bcr^T=0$ the post-processing step reduces the percentage of background $\bcr > \bcr^T$ from $60\%$ to $25\%$ (a $58\%$ improvement) in the August 13 - 21, 2017 time-frame of data. For the entirety of O2, PSD marginalization reduces the percentage of $\bcr > \bcr^T$ from $64\%$ to $33\%$ (a $\sim49\%$ improvement).

### 2022-08-03 titlepage_mnras.tex
Final version

% % Title of the paper, and the short title which is used in the headers.
% % Keep the title short and informative.
% \title[MNRAS \LaTeX\ guide for authors]{\textit{Monthly Notices of the Royal Astronomical
%   Society}:  \LaTeX\ guide for authors}
\title[An IMBH search in O2 using the BCR]{A search for intermediate-mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio}

>>>>

% % Title of the paper, and the short title which is used in the headers.
% % Keep the title short and informative.
% \title[MNRAS \LaTeX\ guide for authors]{\textit{Monthly Notices of the Royal Astronomical
%   Society}:  \LaTeX\ guide for authors}
\title[An IMBH candidate follow-up in O2 using the BCR]{A follow-up on intermediate-mass black hole candidates in the second LIGO--Virgo observing run with the Bayes Coherence Ratio}
